Manuscript_draft

### 2021-03-25 manuscript_draft.Rmd
Work on manny

Growth dominance in a stand considers the contribution of individual trees to total stand growth as a function of their size [@Binkley2006; @Forrester2019; @West2014].  The growth dominance statistic (GDS) expresses growth dominance as a single metric, where a GDS value of 0 indicates a stand where the growth of individual trees, expressed as a proportion of stand total growth, is equal to their size, expressed as a proportion of stand total size (e.g., basal area).  A negative (positive) GDS indicates a stand where smaller (larger) trees contribute more to stand growth relative to their proportional size.  For more information on the formula used to compute GDS, we refer the reader to West [-@West2014].

@@

In 1994, the low RBA unit had 
`r standFitM.contrasts %>%   filter(Contrast=="TU:10 - TU:20") %>%  filter(Variable=="Basal area") %>%  filter(Year==1994) %>%  pull(estimate) %>%   abs() %>%   round(.,1)`
and
 `r standFitM.contrasts %>%   filter(Contrast=="TU:10 - TU:Control") %>% filter(Variable=="Basal area") %>%   filter(Year==1994) %>%   pull(estimate) %>%   abs() %>%   round(.,1)`
 less basal area per hectare than the high RBA unit and control, respectively, and by 2019, those differences had decreased to 
`r standFitM.contrasts %>%   filter(Contrast=="TU:10 - TU:20") %>%  filter(Variable=="Basal area") %>%  filter(Year==2019) %>%  pull(estimate) %>%   abs() %>%   round(.,1)`
and
`r standFitM.contrasts %>%   filter(Contrast=="TU:10 - TU:Control") %>% filter(Variable=="Basal area") %>%   filter(Year==2019) %>%   pull(estimate) %>%   abs() %>%   round(.,1)`, respectively.  In terms of stand density, the low RBA unit accrued `r standFit.contrasts %>% as.data.frame() %>% filter(Variable=="Stand density") %>%   pull(estimate) %>% round(.,0) %>% .[1]` 
and
`r standFit.contrasts %>% as.data.frame() %>% filter(Variable=="Stand density") %>%   pull(estimate) %>% round(.,0) %>% .[2]` more trees per hectare from 1994-2019, than the high RBA and control, respectively (p<0.05for both), and the high RBA accrued  `r standFit.contrasts %>% as.data.frame() %>% filter(Variable=="Stand density") %>%   pull(estimate) %>% round(.,0) %>% .[3]` more trees than the control (p<0.05) over the same time period.  By 2019, there were no significant stand density differences between the treatments and control.

The rate of change varied by measurement period.  In the low RBA unit, basal area and volume increment peaked at 
`r perInc.X %>% filter(BA.Target=="TU:10") %>% filter(Period=="1997-2009") %>% pull("BA") %>%   round(.,2)` m2/ha and
 `r perInc.X %>% filter(BA.Target=="TU:10") %>% filter(Period=="1997-2009") %>% pull("Vol") %>%   round(.,1)` m3/ha, respectively, in the 1997-2009 period; increments declined to `r perInc.X %>% filter(BA.Target=="TU:10") %>% filter(Period=="2009-2019") %>% pull("BA") %>%   round(.,2)` m2/ha and
 `r perInc.X %>% filter(BA.Target=="TU:10") %>% filter(Period=="2009-2019") %>% pull("Vol") %>%   round(.,1)` m3/ha respectively in the subsequent measurement period of 2009-2019.  In the high RBA unit, periodic annual basal area and volume increment increased with each measurement period and peaked in the 2009-2019 measurement period at `r perInc.X %>% filter(BA.Target=="TU:20") %>% filter(Period=="2009-2019") %>% pull("BA") %>%   round(.,2)` m2/ha and
 `r perInc.X %>% filter(BA.Target=="TU:20") %>% filter(Period=="2009-2019") %>% pull("Vol") %>%   round(.,1)` m3/ha respectively.  Conversely, basal area and volume increment in the control unit was highest in the 1994-1997 period, and declined with each subsequent measurement period to `r perInc.X %>% filter(BA.Target=="TU:Control") %>% filter(Period=="2009-2019") %>% pull("BA") %>%   round(.,2)` m2/ha and
 `r perInc.X %>% filter(BA.Target=="TU:Control") %>% filter(Period=="2009-2019") %>% pull("Vol") %>%   round(.,1)` m3/ha respectively..

>>>>

We calculated growth dominance statistic (GDS) for each plot and time period.  Growth dominance in a stand considers the contribution of individual trees to total stand growth as a function of their size [@Binkley2006; @Forrester2019; @West2014].  The GDS expresses growth dominance as a single metric, where a value of 0 indicates a stand where the growth of individual trees, expressed as a proportion of stand total growth (in this case, basal area increment), is equal to their size, expressed as a proportion of stand total size (in this case, basal area).  A negative (positive) GDS indicates a stand where smaller (larger) trees contribute more to stand growth relative to their proportional size.  For more information on the formula used to compute GDS, we refer the reader to West [-@West2014].

@@

In 1994, the low RBA unit had 
`r standFitM.contrasts %>%   filter(Contrast=="TU:10 - TU:20") %>%  filter(Variable=="Basal area") %>%  filter(Year==1994) %>%  pull(estimate) %>%   abs() %>%   round(.,1)`
and
 `r standFitM.contrasts %>%   filter(Contrast=="TU:10 - TU:Control") %>% filter(Variable=="Basal area") %>%   filter(Year==1994) %>%   pull(estimate) %>%   abs() %>%   round(.,1)`
 less basal area (m² per hectare) than the high RBA unit and control, respectively, and by 2019, those differences had decreased to 
`r standFitM.contrasts %>%   filter(Contrast=="TU:10 - TU:20") %>%  filter(Variable=="Basal area") %>%  filter(Year==2019) %>%  pull(estimate) %>%   abs() %>%   round(.,1)`
and
`r standFitM.contrasts %>%   filter(Contrast=="TU:10 - TU:Control") %>% filter(Variable=="Basal area") %>%   filter(Year==2019) %>%   pull(estimate) %>%   abs() %>%   round(.,1)` m² per hectare, respectively.  In terms of stand density, the low RBA unit accrued `r standFit.contrasts %>% as.data.frame() %>% filter(Variable=="Stand density") %>%   pull(estimate) %>% round(.,0) %>% .[1]` 
and
`r standFit.contrasts %>% as.data.frame() %>% filter(Variable=="Stand density") %>%   pull(estimate) %>% round(.,0) %>% .[2]` more trees per hectare from 1994-2019, than the high RBA and control, respectively (p<0.05 for both), and the high RBA accrued  `r standFit.contrasts %>% as.data.frame() %>% filter(Variable=="Stand density") %>%   pull(estimate) %>% round(.,0) %>% .[3]` more trees than the control (p<0.05) over the same time period.  By 2019, there were no significant stand density differences between the treatments and control.

The rate of change varied by measurement period.  In the low RBA unit, basal area and volume increment peaked at 
`r perInc.X %>% filter(BA.Target=="TU:10") %>% filter(Period=="1997-2009") %>% pull("BA") %>%   round(.,2)` m²/ha and
 `r perInc.X %>% filter(BA.Target=="TU:10") %>% filter(Period=="1997-2009") %>% pull("Vol") %>%   round(.,1)` m³/ha, respectively, in the 1997-2009 period; increments declined to `r perInc.X %>% filter(BA.Target=="TU:10") %>% filter(Period=="2009-2019") %>% pull("BA") %>%   round(.,2)` m²/ha and
 `r perInc.X %>% filter(BA.Target=="TU:10") %>% filter(Period=="2009-2019") %>% pull("Vol") %>%   round(.,1)` m³/ha respectively in the subsequent measurement period of 2009-2019.  In the high RBA unit, periodic annual basal area and volume increment increased with each measurement period and peaked in the 2009-2019 measurement period at `r perInc.X %>% filter(BA.Target=="TU:20") %>% filter(Period=="2009-2019") %>% pull("BA") %>%   round(.,2)` m²/ha and
 `r perInc.X %>% filter(BA.Target=="TU:20") %>% filter(Period=="2009-2019") %>% pull("Vol") %>%   round(.,1)` m³/ha respectively.  Conversely, basal area and volume increment in the control unit was highest in the 1994-1997 period, and declined with each subsequent measurement period to `r perInc.X %>% filter(BA.Target=="TU:Control") %>% filter(Period=="2009-2019") %>% pull("BA") %>%   round(.,2)` m²/ha and
 `r perInc.X %>% filter(BA.Target=="TU:Control") %>% filter(Period=="2009-2019") %>% pull("Vol") %>%   round(.,1)` m³/ha respectively..
